Relational programming in Smalltalk

### 2019-03-18 booklet/article.tex
Little fixes to compile with make.

% ********************************************************************* 
% Revisions and comments Macros
% ********************************************************************* 
\usepackage{xcolor}
\usepackage[normalem]{ulem} 			% \sout macro
\newboolean{showcomments}
% \setboolean{showcomments}{false}
\setboolean{showcomments}{true}

\ifthenelse{\boolean{showcomments}}
{
	\newcommand{\nb}[3]{
		{\colorbox{#2}{\bfseries\sffamily\scriptsize\textcolor{white}{#1}}}
		{\textcolor{#2}{\sf\small$\blacktriangleright$\textit{#3}$\blacktriangleleft$}}}
	\newcommand{\bnote}[2]{\fbox{\color{blue}\bfseries\sffamily\scriptsize#1}
    	{\color{blue}\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}}
	\newcommand{\old}[1]{{\color{gray}\sout{#1}}} % old to be removed
	\newcommand{\del}[1]{\old{#1}} % please remove
	\newcommand{\ins}[1]{{\textcolor{blue}{\uline{#1}}}} % please insert	
	\newcommand{\ugh}[1]{{\textcolor{red}{\uwave{#1}}}} % please rephrase	
	\newcommand{\chg}[2]{{\textcolor{red}{\sout{#1}}}{\ra}\textcolor{blue}{\uline{#2}}} % please change
	\newcommand{\here}{\bnote{***}{CONTINUE HERE}} 
	\newcommand{\fix}[1]{\bnote{FIX}{#1}}
}{
	\newcommand{\bnote}[2]{}
	\newcommand{\nb}[3]{}
	\newcommand{\old}[1]{}
	\newcommand{\del}[1]{}
	\newcommand{\ins}[1]{}
	\newcommand{\ugh}[1]{}
	\newcommand{\chg}[2]{}
	\newcommand{\here}{}
	\newcommand{\fix}[1]{}
} 
%---% add your own macros 
\newcommand{\sd}[1]{\nb{Stef}{orange}{#1}}

@@

\s{I do not get fromObject: implementation. because you do not wrap the object.
So 1 asCons will return 1}
and (ii)~the abstract class \ct{Chain} provides the following class-side messages
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Chain class >> bottom
    ^ Bottom new

@@

On the other hand, when a \ct{Chain} object (ii)~receives the message
\ct{#contents} it builds a \ct{LinkedList} object \textit{eagerly},
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Bottom >> contents
    ^ LinkedList new

Knot >> contents
    ^ self next contents
        addFirst: item;
        yourself
\end{minted}
where \Verb|Knot| objects provide the following simple but fundamental message
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Knot>>#next
    ^ linker value
\end{minted}
Observe that the \ct{|Chain} hierarchy supports the protocol provided
by \ct{ReadStream} objects because of the messages \ct{#next}, \ct{#next:}
and \ct{#contents}.

@@

Recall \sd{it was not said that way so I do not recall} that a \ct{Goal} object responds with a \ct{Chain} that contains either zero,
one, or more substitutions when it receives the message \ct{#onState:} and it
delegates the task to process those substitutions to \ct{Runner} objects,

>>>>

% ********************************************************************* 
% Revisions and comments Macros
% ********************************************************************* 
\usepackage{xcolor}
\usepackage{xspace}
\usepackage[normalem]{ulem} 			% \sout macro
\newboolean{showcomments}
% \setboolean{showcomments}{false}
\setboolean{showcomments}{true}

\ifthenelse{\boolean{showcomments}}
{
	\newcommand{\nb}[3]{
		{\colorbox{#2}{\bfseries\sffamily\scriptsize\textcolor{white}{#1}}}
		{\textcolor{#2}{\sf\small$\blacktriangleright$\textit{#3}$\blacktriangleleft$}}}
	\newcommand{\bnote}[2]{\fbox{\color{blue}\bfseries\sffamily\scriptsize#1}
    	{\color{blue}\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}}
	\newcommand{\old}[1]{{\color{gray}\sout{#1}}} % old to be removed
	\newcommand{\del}[1]{\old{#1}} % please remove
	\newcommand{\ins}[1]{{\textcolor{blue}{\uline{#1}}}} % please insert	
	\newcommand{\ugh}[1]{{\textcolor{red}{\uwave{#1}}}} % please rephrase	
	\newcommand{\chg}[2]{{\textcolor{red}{\sout{#1}}}{\ra}\textcolor{blue}{\uline{#2}}} % please change
	\newcommand{\here}{\bnote{***}{CONTINUE HERE}} 
	\newcommand{\fix}[1]{\bnote{FIX}{#1}}
}{
	\newcommand{\bnote}[2]{}
	\newcommand{\nb}[3]{}
	\newcommand{\old}[1]{}
	\newcommand{\del}[1]{}
	\newcommand{\ins}[1]{}
	\newcommand{\ugh}[1]{}
	\newcommand{\chg}[2]{}
	\newcommand{\here}{}
	\newcommand{\fix}[1]{}
} 
%---% add your own macros 
\newcommand{\sd}[1]{\nb{Stef}{orange}{#1}}
\newcommand{\mn}[1]{\nb{Massimo}{blue}{#1}}

@@

\sd{I do not get fromObject: implementation. because you do not wrap the object.
So 1 asCons will return 1}
and (ii)~the abstract class \ct{Chain} provides the following class-side messages
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Chain class >> bottom
    ^ Bottom new

@@

On the other hand, when a \ct{Chain} object (ii)~receives the message
\ct{\#contents} it builds a \ct{LinkedList} object \textit{eagerly},
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Bottom >> contents
    ^ LinkedList new

Knot >> contents
    ^ self next contents
        addFirst: item;
        yourself
\end{minted}
where \Verb|Knot| objects provide the following simple but fundamental message
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Knot>>#next
    ^ linker value
\end{minted}
Observe that the \ct{Chain} hierarchy supports the protocol provided
by \ct{ReadStream} objects because of the messages \ct{\#next}, \ct{\#next:}
and \ct{\#contents}.

@@

Recall \sd{it was not said that way so I do not recall} that a \ct{Goal} object responds with a \ct{Chain} that contains either zero,
one, or more substitutions when it receives the message \ct{\#onState:} and it
delegates the task to process those substitutions to \ct{Runner} objects,
